---
title: "eswa"
author: "Esther_Wairimu_Kamau"
date: "1/10/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include= TRUE}
library(knitr)
library(data.table)
library(tidyverse)
library(tigerstats)
tinytex::install_tinytex() 
opts_chunk$set(echo = FALSE)
```

# Defining the problem
The research problem in this case is to find out individuals that are likely to click on a blog advert based on their characteristics which include; Age Daily Time spent on site Area of residence Internet Usage Gender Country of residence

# Metric of Success

The metric success of this project is to identify clients likely to click on the ad after performing intense data analysis(EDA).

# Data Relevance
The data provided by the client is from the performance of a previous blog advert on the same website. 
The columns are as follows:
* **Daily Time Spent on the site-Integer**
* **Age of the individual browsing-Integer**
* **Area of residence Internet Usage** 
* **Gender of the browsing individual**
* **Country of Residence**

# Understanding the Context

# Experimental Design
*.Data Loading 
*.Data cleaning for missing values and outliers
*.Exploratory Data Analysis 
*.Conclusion-Detecting the trend in behaviour.

# Data Loading and exploring

### Importing our dataset


```{r}
advertising =read.csv('http://bit.ly/IPAdvertisingData',header = TRUE, sep = ",",fileEncoding = "UTF-8-BOM")

```

### exploring the top of our data

```{r}
head(advertising)
```


### exploring the bottom of our data
```{r}
tail(advertising)
```



### Checking the class of the object "advertising"

```{r}
class(advertising)
```


### Checking the dimension of our dataset

```{r}
dim(advertising)
```



### Checking the structure of our data frame


```{r}
str(advertising)
```



### Getting the names of the columns we will be working with

```{r}
colnames(advertising)
```



# Data cleaning

### Checking for duplicated values in our data set
```{r}
anyDuplicated(advertising)
```


### Checking if our dataset has any missing values

```{r}
sum(is.na(advertising))
```

### checking for missing values using case.complete function(just to confirm)

```{r}
# The function complete.cases() returns a logical vector indicating which cases are complete.
# list rows of data that have missing values
advertising[!complete.cases(advertising),]
```


### checking for outliers, we only need the numerical columns

```{r}
nums <- unlist(lapply(advertising, is.numeric)) 

numerical_cols <- advertising[ ,nums]

head(numerical_cols)
```


### checking for unique values

```{r}
unique_items <- unique(advertising)
unique_items
```



# Plotting the boxplot to visualize the outliers in the dataset

```{r}
boxplot(numerical_cols[,-1], horizontal=TRUE, main="Ad campaign outliers")

```


# Univariate Analysis

## For ease in analysis,we convert the data into a tibble

```{r}
library(tidyverse)

my_data<-as_tibble(advertising)
my_data

```


## Numerical data

```{r}
library(dplyr)
Numeric=my_data %>% select_if(is.numeric)
Numeric
```


## categorical data
```{r}
Categorical=my_data %>% select_if(is.character)
Categorical
```

## find the descriptive statistics of the numerical data

```{r}
summary(Numeric)
```

# Bivariate and Multivariate Analysis

## comparing covariances

## checking correlations

## Scatter Plots

## pair plots

# Observations

# conclusions

# Recommendations



# 

